WHOOMP! (There It Is) \\ Rapid Bayesian position reconstruction for gravitational\nobreakdashes-wave transients

### 2015-07-07 ms.tex
Refer to autocorrelation likelihood figure
Rework acknowledgements

A more natural way is to plug the signal model from Equation~(\ref{eq:signal-model}) \emph{evaluated at the \ac{ML} parameter estimates} into the single-detector likelihood in Equation~(\ref{eq:gaussian-likelihood-spa}):
%
\begin{widetext}
\begin{equation}
    p\left(\hat{\bm\theta}_i \middle| \bm\theta \right)
    :=
    p\left(Y_i(\omega) = X_i(\omega; \hat{\bm\theta})
        \middle| \bm\theta \right)
    %\\
    \propto \exp \Bigg[
        - \frac{1}{2} \int_0^\infty \Bigg|
            \frac{\hat{\rho}_i}{\sigma_i(\hat{\bm\theta}_\mathrm{in})} e^{i (\hat\gamma_i - \omega \hat\tau_i)} \frac{H(\omega; \hat{\bm\theta}_\mathrm{in})}{S_i(\omega)}
    %\\
            - \frac{\rho_i}{\sigma_i(\bm\theta_\mathrm{in})} e^{i (\gamma_i - \omega \tau_i)} \frac{H(\omega; \bm\theta_\mathrm{in})}{S_i(\omega)}
        \Bigg|^2 \, d\omega
    \Bigg].
\end{equation}
\end{widetext}
%
If we further assume that \emph{the intrinsic parameters are equal to their \ac{ML} estimates}, $\bm\theta_\mathrm{in} = \hat{\bm\theta}_\mathrm{in}$, then this reduces to what we call the \emph{autocorrelation likelihood},
%
\begin{widetext}
\begin{equation}\label{eq:autocor-likelihood}
    p\left(\hat\rho_i, \hat\gamma_i, \hat\tau_i
        \middle| \rho_i, \gamma_i, \tau_i \right) \propto
    \exp \left[ - \frac{1}{2}{\hat\rho_i}^2 - \frac{1}{2}{\rho_i}^2
        + \hat\rho_i \rho_i \Re \left\{ e^{i \tilde{\gamma}_i} a_i^*(\tilde{\tau}_i)
        \right\}
    \right],
\end{equation}
\end{widetext}
%
with $\tilde{\gamma}_i = \hat\gamma_i - \gamma_i$, $\tilde{\tau}_i = \hat\tau_i - \tau_i$, and the template's autocorrelation function $a_i(t; \bm\theta_\mathrm{in})$ defined as
%
\begin{equation}\label{eq:autocorrelation-function}
    a_i(t; \bm\theta_\mathrm{in}) := \frac{1}{{{\sigma_i}^2(\bm\theta_\mathrm{in})}} \int_0^\infty \frac{\left| H(\omega; \hat{\bm\theta}_\mathrm{in})\right|^2}{S_i(\omega)} e^{i \omega t} \,d\omega.
\end{equation}
%
To assemble the joint likelihood for the whole network, we form the product of the autocorrelation likelihoods from the individual detectors:
%
\begin{multline}
    p\left(\left\{\hat\rho_i, \hat\gamma_i, \hat\tau_i\right\}_i
        \middle| \left\{\rho_i, \gamma_i, \tau_i\right\}_i \right) \\
    \propto
    \exp \left[ - \frac{1}{2} \sum_i {\hat\rho_i}^2 - \frac{1}{2} \sum_i {\rho_i}^2
        + \sum_i \hat\rho_i \rho_i \Re \left\{ e^{i \tilde{\gamma}_i} a^*(\tilde{\tau}_i)
        \right\}
    \right].
\end{multline}

@@

\begin{acknowledgements}
Source code for \ac{BAYESTAR} is available as part of LALInference, the open source LIGO/Virgo parameter estimation toolchain. See \url{http://www.lsc-group.phys.uwm.edu/daswg/projects/lalsuite.html} for further details.

@@

LIGO was constructed by the California Institute of Technology and Massachusetts Institute of Technology with funding from the \ac{NSF} and operates under cooperative agreement PHY\nobreakdashes-0107417. Some results were produced on the NEMO computing cluster operated by the Center for Gravitation and Cosmology at University of Wisconsin\nobreakdashes--Milwaukee under \ac{NSF} Grants PHY\nobreakdashes-0923409 and PHY\nobreakdashes-0600953. This research is supported by the \ac{NSF} through a Graduate Research Fellowship to L.S. This paper has LIGO Document Number LIGO\nobreakdashes-P1500009\nobreakdashes-v1.

>>>>

A more natural way is to plug the signal model from Equation~(\ref{eq:signal-model}) \emph{evaluated at the \ac{ML} parameter estimates} into the single-detector likelihood in Equation~(\ref{eq:gaussian-likelihood-spa}):
%
\begin{widetext}
\begin{equation}
    p\left(\hat{\bm\theta}_i \middle| \bm\theta \right)
    :=
    p\left(Y_i(\omega) = X_i(\omega; \hat{\bm\theta})
        \middle| \bm\theta \right)
    %\\
    \propto \exp \Bigg[
        - \frac{1}{2} \int_0^\infty \Bigg|
            \frac{\hat{\rho}_i}{\sigma_i(\hat{\bm\theta}_\mathrm{in})} e^{i (\hat\gamma_i - \omega \hat\tau_i)} \frac{H(\omega; \hat{\bm\theta}_\mathrm{in})}{S_i(\omega)}
    %\\
            - \frac{\rho_i}{\sigma_i(\bm\theta_\mathrm{in})} e^{i (\gamma_i - \omega \tau_i)} \frac{H(\omega; \bm\theta_\mathrm{in})}{S_i(\omega)}
        \Bigg|^2 \, d\omega
    \Bigg].
\end{equation}
\end{widetext}
%
If we further assume that \emph{the intrinsic parameters are equal to their \ac{ML} estimates}, $\bm\theta_\mathrm{in} = \hat{\bm\theta}_\mathrm{in}$, then this reduces to what we call the \emph{autocorrelation likelihood},
%
\begin{widetext}
\begin{equation}\label{eq:autocor-likelihood}
    p\left(\hat\rho_i, \hat\gamma_i, \hat\tau_i
        \middle| \rho_i, \gamma_i, \tau_i \right) \propto
    \exp \left[ - \frac{1}{2}{\hat\rho_i}^2 - \frac{1}{2}{\rho_i}^2
        + \hat\rho_i \rho_i \Re \left\{ e^{i \tilde{\gamma}_i} a_i^*(\tilde{\tau}_i)
        \right\}
    \right],
\end{equation}
\end{widetext}
%
with $\tilde{\gamma}_i = \hat\gamma_i - \gamma_i$, $\tilde{\tau}_i = \hat\tau_i - \tau_i$, and the template's autocorrelation function $a_i(t; \bm\theta_\mathrm{in})$ defined as
%
\begin{equation}\label{eq:autocorrelation-function}
    a_i(t; \bm\theta_\mathrm{in}) := \frac{1}{{{\sigma_i}^2(\bm\theta_\mathrm{in})}} \int_0^\infty \frac{\left| H(\omega; \hat{\bm\theta}_\mathrm{in})\right|^2}{S_i(\omega)} e^{i \omega t} \,d\omega.
\end{equation}
%
Some example autocorrelation functions and corresponding likelihoods are shown in Figure~\ref{fig:autocorr-likelihood}. To assemble the joint likelihood for the whole network, we form the product of the autocorrelation likelihoods from the individual detectors:
%
\begin{multline}
    p\left(\left\{\hat\rho_i, \hat\gamma_i, \hat\tau_i\right\}_i
        \middle| \left\{\rho_i, \gamma_i, \tau_i\right\}_i \right) \\
    \propto
    \exp \left[ - \frac{1}{2} \sum_i {\hat\rho_i}^2 - \frac{1}{2} \sum_i {\rho_i}^2
        + \sum_i \hat\rho_i \rho_i \Re \left\{ e^{i \tilde{\gamma}_i} a^*(\tilde{\tau}_i)
        \right\}
    \right].
\end{multline}

@@

\begin{acknowledgements}
This paper has LIGO Document Number LIGO\nobreakdashes-P1500009\nobreakdashes-v1.

LIGO was constructed by the California Institute of Technology and Massachusetts Institute of Technology with funding from the \ac{NSF} and operates under cooperative agreement PHY\nobreakdashes-0107417.

This research is supported by the \ac{NSF} through a Graduate Research Fellowship to L.S.

L.S. thanks the Aspen Center for Physics and \ac{NSF} Grant \#1066293 for hospitality during the editing of this paper.

Source code for \ac{BAYESTAR} is available as part of LALInference, the open source LIGO/Virgo parameter estimation toolchain. See \url{http://www.lsc-group.phys.uwm.edu/daswg/projects/lalsuite.html} for further details.

@@

Some results were produced on the NEMO computing cluster operated by the Center for Gravitation and Cosmology at University of Wisconsin\nobreakdashes--Milwaukee under \ac{NSF} Grants PHY\nobreakdashes-0923409 and PHY\nobreakdashes-0600953.
